### 2023-03-15 chapter0/chapter0.tex
Added aims overview figure

These comparisons, however, require the identification of IDRs with common ancestry that perform ``equivalent'' functions across many distinct organisms. Given the difficulties with identifying similar IDRs by their sequences, this may seem like a chicken and egg problem. Fortunately, IDRs are frequently associated with more conserved folded domains. Thus, identifying evolutionarily related proteins by their overall sequence signatures and ``aligning'' them will in turn identify the equivalent IDRs in those sequences. The first step of an evolutionary analysis of IDRs, then, is the identification of proteins with common ancestry, called orthologs. Since the first genomes were sequenced in the late 1990s, researchers have developed techniques for identifying and aligning orthologs~\cite{Fleischmann1995, Goffeau1996, CESC1998, Tatusov1997}. While these methods are generally effective, they are conducted by automated computational pipelines and prone to errors when processing the highly divergent sequences that characterize many IDRs. The evolutionary relationships between the genomes of closely related species generally make such mistakes easier to identify, and fortunately over the past five years advances in DNA sequencing technology have yielded dramatic increases in the number of sequenced genomes in the \textit{Drosophila} genus. However, because the existing methods for ortholog identification were designed for fewer or more distantly related genomes, they do not fully leveraged the available genomic redundancy to minimize errors. Thus, in the first chapter I develop a novel method for identifying orthologs which addresses this shortcoming and apply it to 33 \textit{Drosophila} genomes to generate a set of aligned orthologs. In the second chapter I then identify rapidly evolving IDRs in these ``alignments'' and analyse them with a variety of evolutionary models to detect patterns of conservation. Finally, in the third chapter I discuss several software tools and tutorials for fitting statistical models to data, which were created while pursuing the previous aims.

>>>>

\begin{figure}[h!]
\includegraphics[width=\textwidth]{out/aims_overview.png}
\centering
\caption{\textbf{Graphical overview of aims.}
\textbf{(A)} The large open circles represent the genomes of different \textit{Drosophila} species, and the small, filled circles represent protein sequences in those genomes. Orthologs are colored with shades of the same hue. \textbf{(B-D)} Each horizontal line represents a sequence in panel A, and together each set of lines creates an ``alignment.'' The lines are broken where ``gaps'' are inserted to align equivalent segments. The grey segments denote regions of the alignment which are not part of subsequent analyses. \textbf{(E)} Each aligned IDR is scored on four features, where the ``strength'' of that feature is indicated by the height and color of its bar. \textbf{(F)} Scoring many IDRs on these features yields three distinct clusters.}
\label{fig:aims_overview}
\end{figure}

These comparisons, however, require the identification of IDRs with common ancestry that perform ``equivalent'' functions across many distinct organisms. Given the difficulties with identifying similar IDRs by their sequences, this may seem like a chicken and egg problem. Fortunately, IDRs are frequently associated with more conserved folded domains. Thus, identifying evolutionarily related proteins by their overall sequence signatures and ``aligning'' them will in turn identify the equivalent IDRs in those sequences. The first step of an evolutionary analysis of IDRs, then, is the identification of proteins with common ancestry, called orthologs. Since the first genomes were sequenced in the late 1990s, researchers have developed techniques for identifying and aligning orthologs~\cite{Fleischmann1995, Goffeau1996, CESC1998, Tatusov1997}. While these methods are generally effective, they are conducted by automated computational pipelines and prone to errors when processing the highly divergent sequences that characterize many IDRs. The evolutionary relationships between the genomes of closely related species generally make such mistakes easier to identify, and fortunately over the past five years advances in DNA sequencing technology have yielded dramatic increases in the number of sequenced genomes in the \textit{Drosophila} genus. However, because the existing methods for ortholog identification were designed for fewer or more distantly related genomes, they do not fully leveraged the available genomic redundancy to minimize errors. Thus, in the first chapter I develop a novel method for identifying orthologs which addresses this shortcoming and apply it to 33 \textit{Drosophila} genomes to generate a set of aligned orthologs (Fig.~\ref{fig:aims_overview}A-B). In the second chapter I then identify rapidly evolving IDRs in these ``alignments'' and analyse them with a variety of evolutionary models to detect patterns of conservation (Fig.~\ref{fig:aims_overview}C-F). Finally, in the third chapter I discuss several software tools and tutorials for fitting statistical models to data, which were created while pursuing the previous aims.
